Accelerated evolution of convective simulations

### 2018-06-15 referee_process/first_submission/referee_response.tex

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[aps, 11pt, singlecolumn]{revtex4-1} % Set the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\usepackage{natbib}
\bibliographystyle{apsrev}
\usepackage{setspace}

\begin{document}
\newenvironment{myquotation}{
\begin{quotation}
\itshape
}{ 
\end{quotation}
}
%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
\noindent
Dear Eckart, other PRFluids editors, and anonymous referees,
$\,$\newline

\begin{singlespace}
We return our revised manuscript, ``Accelerated evolution in
convective simulations,'' addressing the concerns of the referee reports.
We would like to point out a few major changes in this version of the manuscript.
A brief new section, ``Computational time-savings of AE'' has been added after the
results section and before the conclusion section. In this section, we directly
compare the computational cost of AE and SE for select simulations and we discuss
the computational benefits of AE.  Furthermore, the final section of our paper
is now titled ``Discussion \& Conclusions;'' we have removed a vast amount
of speculation regarding extensions of AE from this section and have added
a brief paragraph containing some caveats regarding challenges that users will face in
extending AE to more complex systems.  In addition to these changes in the text,
we have added a new Fig. 2 which compares the evolution of Nu vs time to the
evolution of the
mean temperature vs. time in AE and SE, 
and we have added some accompanying text to the results section for this
figure.  Small changes in the text throughout the paper have
been made where these changes were requested by the reports.

@@

The authors come from an astrophysical convection background where the
Peclet number is generally high and where these techniques are useful.
They need to be careful to be clear that the convection they are
talking about is high Pe, the parameter that describes the separation
of overturning and thermal diffusion timescales. These techniques will
not be useful (or required) for low Pe number situations.
\end{myquotation}
We have reworked some text in the second paragraph of the introduction section
in order to clarify that we are discussing high Pe convection.

\begin{myquotation}
Section II.

After (4): The expression for $\Delta T_0$ assumes that the background
temperature gradient is constant, which has not been introduced yet.
\end{myquotation}
We have added a sentence before the description of our nondimensionalization
stating that the initial temperature profile is linear.

\begin{myquotation}
The sentence ``Our choice of the thermal boundary conditions in Eqn.
(9 ) was motivated by the fact that accelerated evolution is simpler
when both the thermal profile and the flux through the domain are
fixed at a boundary'' is never revisited or explained, unless I missed
it?
\end{myquotation}
We have clarified this in the third to last pargraph of section III,
and added a forward reference to section III by this quoted sentence.

@@

\begin{myquotation}
When discussing the PDFs, the author refer to the ``modes" of the PDF?
This sounds a bit odd to me. Do they mean the peaks? The phrase
``temperature fluctuations off of the modes'' is particularly baffling!
\end{myquotation}
We have replaced ``modes'' with ``maxima.''

\begin{myquotation}
The fact that the mean temperature profile is off by a constant factor
almost everywhere is a bit disturbing, even if the factor is small.
Does continued iteration of AE not get rid of this? If not, then this
is significant, because it would take a thermal relaxation time again
to correct this issue.
\end{myquotation}
We have added a paragraph to the ``accelerated evolution recipe'' appendix.
In it, we discuss the fact that AE can be iteratively solved until you converge
to a desired criterion, and that we were only aiming for O(1\%) in this
paper. With a smaller convergence criterion, the user
could get closer to the evolved SE solution, decreasing the separation
between the temperature PDFs seen in Fig. 5 (previously Fig. 4). 
However, smaller convergence criteria
come with significant computational cost (the horizontal averages must be
better converged, and likely more iterative solves must be performed).

\begin{myquotation}
I felt like I wanted a section at the end of this section that
described how much the AE saved in computational time and wall time
over SE. There is a bit of it here, and some of it in the discussion,
and some more details in the appendix, which seems a bit scattered. I
would be tempted to agglomerate it all here in a separate section.
\end{myquotation}
We have added a new section V., ``Computational Time-savings of AE.''
We have combined the relevant information which was previously in the results 
section and the conclusion section here, 
and we have added some information on a few
additional runs in a new table.

@@

We have corrected our error and made our meaning clearer.

\begin{myquotation}
$\,$\\\vspace{-1.25cm}
\begin{enumerate}
\setcounter{enumi}{3}
\item It is true that direct numerical simulation of turbulent thermal convection is expensive. However, the highest Ra simulation run by Stevens et al. reaches a stationary state in a few hundreds of free-fall times. Hence, what is the necessity of running simulations for ``thousands or millions of free-fall times''?
\end{enumerate}
\end{myquotation}
We have added a new Fig. 2 to address this question. In general, we find that
the heat transport reaches its final, time-stationary mean
value very quickly (see panels (a) and (b) of Fig. 2). However, the thermal 
structure, kinetic energy, and fluctuations of heat transport away from the mean
can still slowly evolve over long timescales
despite the flows being relatively statistically stationary. The purpose of
running for many, many free-fall times is to converge the thermal profile of
the domain to ensure that the full system is in a steady state.

\begin{myquotation}
$\,$\\\vspace{-1.25cm}
\begin{enumerate}
\setcounter{enumi}{4}
\item Presumably, the most important step in the “Accelerated convergence method” involves decreasing the heat flux through the top boundary so as to match it with the heat flux at the bottom boundary. This is achieved by introducing a function $\xi(z) \equiv  F_B/F_{tot}$. The following questions arise:
\begin{enumerate}
\item What is the functional form of $\xi(z)$? And where does the z-dependence come from?
\item Here, $F_{tot}$ is not really a constant, but depends on time. Hence, $\xi(z)$ should also be a function of time.
\item The ``evolved'' quantities are obtained by multiplying $\langle u \times \omega \rangle$ and $F_E$ by $\xi(z)$. Can this construction be rigorously justified? I would like the authors to provide more details on how they arrived at this step.
\end{enumerate}
\end{enumerate}
\end{myquotation}

(a) In the simple case studied in this problem, where $F_B$ is constant in time
and space, the functional form of $\xi \propto F_{tot}^{-1}$, and depends only
on the fluxes of the convective system. Since $F_{tot}$ can
have a large z-dependence (e.g., Fig. 1b), $\xi$, too, has a z-dependence. 
$\xi$ is simply a height-dependent scaling that we use to appropriately
adjust our velocity, temperature, etc. We have added a sentence after the 
definition of $\xi$ giving an example of its value for the case presented in
Fig. 1.

@@

(b) We have changed $\xi(z) \rightarrow \xi(z, t)$.

(c) Physically, $\xi$ is a scaling factor that tells us how much larger the flux
is in the interior than it should be. 
We assume that $F_{tot} \sim F_{conv} \sim w (T - \langle T \rangle )$.
As such, if we must scale $F_{tot}$ by a factor of $\xi$, we are equivalently
scaling
$u$ and $T - \langle T\rangle$ by $\sqrt{\xi}$ to reduce the convective flux.
We scale $\langle u\times \omega\rangle$ by $\xi$, as it is
nondimensionally of order $u^2 / L$, and so if the velocities must be multiplied
by $\sqrt{\xi}$, this term must be multiplied by $\xi$. We have added a short
paragraph after Eqns (12) and (13) containing this information.

### 2018-07-19 tex/bvp_paper.tex
referee round 2

\begin{abstract}
High Peclet number, turbulent convection is a classic system with a large timescale
separation between flow speeds and equilibration time.
In this paper, we present a method of achieving Accelerated Evolution (AE) of convective simulations which
fast-forwards through the long thermal evolution of these systems, and we test the
validity of this method. The AE procedure involves measuring the dynamics of convection
early in a simulation and using its characteristics to adjust the mean thermodynamic
profile within the domain towards its evolved state. We study \RB convection as a test case for AE.  
Evolved flow properties of AE solutions are measured to be within a few percent
of solutions which are reached through Standard Evolution (SE) over a full thermal timescale.
At the highest values of the Rayleigh number at which we compare SE and AE,
we find that AE solutions require roughly an order of magnitude fewer cpu-hours
to compute than SE solutions.
\end{abstract}
\maketitle

@@

\section{Introduction}
\label{sec:intro}
Astrophysical convection occurs in the presence of disparate timescales. 
Studying realistic models of natural systems through direct numerical
simulations is infeasible due to the
large separation between various flow timescales and equilibration times.
Stiffness in astrophysical systems can manifest in multiple ways, some of which can
be handled by clever choices of numerical algorithms, and some which cannot.
For example,
flows in the convection zones of stars like the Sun are characteristically low Mach number
(Ma) in the deep interior. Initial value problems solved using
explicit timestepping methods are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit corresponding to the fastest motions (sound
waves), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff, and the difference between
the sound crossing time and the convective overturn time has made studies of low-Ma stellar
convection difficult. This stiffness can be avoided using approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out
\cite{brown&all2010, featherstone&hindman2016}.
Recently, advanced numerical techniques which use fully implicit 
\cite{viallet&all2011, viallet&all2013, viallet&all2016} or mixed
implicit-explicit \cite{lecoanet&all2014, anders&brown2017, bordwell&all2018} 
timestepping mechanisms have made it possible to study
convection in the fully compressible equations at low Mach numbers, 
and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

Unfortunately, astrophysical convective systems are stiff in more than one 
timescale. Specifically, the Peclet number (Pe), the ratio of the thermal
diffusion timescale to the convective velocity timescale, is large.
In a high Pe system, many convective timescales must pass before 
the domain reaches thermal equilibrium, a steady state in which
evolution of the thermal structure of the convective region has ceased.
As the timestep size of implicit methods is bound to the fastest nonlinear flows
(convection), fully implicit methods cannot be used to address this form of stiffness
\cite{viallet&all2011, viallet&all2013, viallet&all2016}.

Resolving dynamics in atmospheres which are sufficiently
thermally equilibrated therefore remains a challenging problem.
Solar convection is a prime example of this phenomenon, as
dynamical timescales in the solar convective zone are relatively short 
(convection overturns every $\sim 5$ minutes at the solar surface)
compared to the Sun's thermal evolution timescale, which is O($10^7$) years
\cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
meaningfully evolving the thermal structure of the system using
traditional timestepping techniques alone.
As modern simulations aim to model natural, high-Pe convection
in the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale ($t_{\kappa}$, defined in section \ref{sec:ae}) 
becomes intractably large compared to dynamical timescales 
such as the freefall time ($t_{\text{ff}}$, defined in section \ref{sec:experiment})
\cite{anders&brown2017}, 
\begin{equation}
\frac{t_{\kappa}}{t_{\text{ff}}} \propto (\text{Ra})^{1/2}.
\end{equation}
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Capturing appropriately resolved
turbulent motions requires finer grid meshes and smaller timesteps.
Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
and the number of overturn times required for systems to reach thermal equilibrium
grows.

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally equilibrated solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the flow
fields in a converged solution at low Ra as initial conditions for a simulation at high
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  In systems in which there are multiple
stable solutions, such as a roll state and a shearing state of convection,
the large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
It is possible that this puts the high Ra solution into a different stable state
than it would naturally reach from an initial, hydrostatically stable configuration.
Another commonly-used tactic in
moderate-Ra simulations is to first solve a simple model of the system in question,
and then use the solution of that model as initial conditions for full nonlinear
direct numerical simulations. For example, past studies have used initial conditions from
a linear eigenvalue solve \cite{hurlburt&all1984} in plane-parallel studies, or
axisymmetric solutions in studies of convection in 3D cylinders \cite{verzicco&camussi1997}. 
In other systems, particularly when convective zones are adjacent to stable regions,
authors often choose initial conditions which are not in a classic, hydrostatic,
conductive state. Rather, either through knowledge of low-Ra solutions \cite{couston&all2017}
or broader convective theories such as Mixing Length Theory
\cite{brandenburg&all2005}, initial conditions can be adjusted such that the stratification
within the convective domain is closer to an equilibrium state than the largely
unstable hydrostatic state.

Despite the numerous methods that have been used,
the most straightforward way to achieve a thermally equilibrated solution
is to evolve a convective simulation through a thermal timescale. Some modern
studies do just that \cite{featherstone&hindman2016}.
Such evolution is computationally 
\emph{expensive}, and state-of-the-art simulations at the highest values of Ra
can only reasonably be run
for hundreds of freefall timescales \cite{stevens&all2011}, much less the
thousands or millions of freefall times required for thermal equilibration.

In this work, we study a method of achieving accelerated evolution of
convective simulations. We couple measurements of the dynamics of unequilibrated
convective simulations with knowledge about energy balances in the desired solution
to self-consistently adjust the mean vertical thermodynamic profile towards its evolved state. 
While a technique of this kind has been used previously \cite{hurlburt&all1986},
the details of implementation, the convergence properties, and whether or not the
accelerated evolution state corresponds to a standard evolution state are not documented.
In section \ref{sec:experiment}, we describe our convective simulations and
numerical methods. In section \ref{sec:ae}, we describe the procedure for 
achieving accelerated evolution. In
section \ref{sec:results}, we compare accelerated evolution solutions
to solutions obtained from standard evolution through a full thermal diffusion timescale. 
In section \ref{sec:speedups}, we compare the numerical cost of accelerated
solutions to standard solutions.
Finally,
in section \ref{sec:extensions}, we offer concluding remarks and
discuss extensions of the methods presented here.

@@

We study in depth a 2D simulation at $S = 10^5$ to demonstrate the power of AE.
We compare kinetic energy (KE, black line) and mean temperature (blue line)
traces from a SE run in Fig.~\ref{fig:time_trace}a to an AE run
in Fig.~\ref{fig:time_trace}c.
In Fig.~\ref{fig:time_trace}a, the time evolution of the SE simulation is shown.
The KE grows exponentially from white noise during the
first $\sim 25$ $t_{\text{ff}}$. The solution then saturates and begins to slowly
equilibrate towards the proper isothermal profile in the interior of the domain.
This slow equilibration is evident in the behavior of the blue line, which measures
$\angles{T} - T_{\text{top}}$, where \angles{T} is the volume-average of $T$, and
$T_{\text{top}} = -0.5$ is the temperature at the upper boundary.
The mean atmospheric temperature and
kinetic energies are fully converged when $t = 4000$ $t_{\text{ff}}$ $= 0.35$ $t_{\kappa}$.
We show roughly the first thousand freefall time
units of evolution, as well as the evolved thermodynamic state reached after a full
thermal time of evolution.  In Fig.~\ref{fig:time_trace}c, similar traces are
shown for the corresponding AE solution
at the same parameters. The same linear growth phase occurs, but shortly after
the peak of convective transient we accelerate the convergence of the atmosphere
through the process which we describe below. We adjust the 1D vertical profile of
the atmosphere three times, as denoted by the three labeled
arrows in the graph numbered 1-3.  The third profile adjustment associated with
arrow 3 is small enough (see appendix \ref{appendix:recipe}) that we assume
the atmosphere is sufficiently converged, and we begin to sample the evolved
convective dynamics.

@@

In order to adjust the temperature profile to achieve AE, we calculate the total flux,
F$_{\text{tot}} =$ F$_{\text{E}}$ + F$_{\kappa}$, and then derive the profiles
\vspace{-0.5cm}
\begin{equation}
f_{\text{E}}(z) = \frac{\text{F}_{\text{E}}}{\text{F}_{\text{tot}}},\qquad
f_{\kappa}(z) = \frac{\text{F}_{\kappa}}{\text{F}_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
which have the systematic asymmetries (Fig.~\ref{fig:time_trace}b) removed. These profiles describe which
parts of the atmosphere depend on convection to carry flux (where $f_{\text{E}}(z) = 1$
and $f_{\kappa}(z) = 0$).
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the extent of the early thermal boundary layers 
(where $f_{\kappa}(z) = 1$ and $f_{\text{E}}(z) = 0$) 
will not change significantly over the course of the atmosphere's evolution.
Under this assumption, in order to reach the converged state, 
the flux through the atmosphere must be decreased by some amount,
\vspace{-0.5cm}
\begin{equation}
\xi(z,t) \equiv \frac{\text{F}_{\text{B}}}{\text{F}_{\text{tot}}},
\label{eqn:xi}
\end{equation}
where $\text{F}_{\text{B}} = \mathcal{P}$ is the amount of flux that enters the
atmosphere at the bottom, fixed-flux boundary.
For example, in Fig.~\ref{fig:time_trace}b,
F$_{\text{tot}} \approx 19\mathcal{P}$ at the upper boundary,
but in the equilibrated state 
it should just be $\mathcal{P}$, so $\xi \approx 1/19$ at that depth.

@@

\begin{figure}[b]
\includegraphics[width=\textwidth]{./figs/nu_v_time.png}
\caption{The time evolution of Nu (black) and $\angles{T} - T_{\text{top}}$ (blue)
are shown for simulations at $S = 10^5$ in SE (a) and AE (b).
A moving average of Nu, using a centered boxcar with a span of 50 freefall time
units, is overplotted as a grey line. 
We see that the mean value of Nu reaches its equilibrium value quickly
compared to $\angles{T}$.
Regardless, fluctuations of Nu about the mean
value are much smaller in the thermally equilibrated state.
The times at which AE solves occur are denoted by arrows in (b),
and the grey boxed region is examined in more detail in Fig. \ref{fig:oscillating_plumes}.
\label{fig:nu_v_time} }
\end{figure}

@@

The Nusselt number (Nu) quantifies the efficiency of convective heat transport
and is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}},
\end{equation}
where the volume average of a quantity, $\eta$, is shown as $\angles{\eta}$.
The time evolution of Nu in AE and SE is compared to the mean temperature evolution
in Fig.~\ref{fig:nu_v_time}. In Fig.~\ref{fig:nu_v_time}a, we show the evolution
of the SE run at $S = 10^5$. The black trace shows the instantaneous value of
Nu, and the overplotted grey line shows a moving time average of Nu. The
time average is taken using a centered boxcar window whose width is 50 $t_{\text{ff}}$.
The mean value of Nu evolves towards its final value (dashed horizontal
grey line) rapidly compared to $\angles{T}-T_{\text{top}}$ (blue), 
which approaches its evolved state (dashed horizontal blue line) slowly.
As such, it is possible to measure the statistically stationary mean value of
Nu after only a few hundred simulated freefall times, 
as has been done previously \cite{stevens&all2010}. However, fluctuations in
Nu around the mean value at early times have both a larger magnitude and higher
frequency than the fluctuations in the equilibrated state.
In Fig.~\ref{fig:nu_v_time}b, we show the AE traces of Nu and
$\angles{T}-T_{\text{top}}$ at $S = 10^5$. The times at which AE solves occur are marked
by arrows. Nu reaches its evolved mean value slightly more rapidly than in the SE case,
and the frequency and magnitude of fluctuations in Nu away from the mean resemble the
final equilibrated state of SE.

>>>>

\begin{abstract}
High Peclet number, turbulent convection is a classic system with a large timescale
separation between flow speeds and the thermal relaxation time.
In this paper, we present a method of achieving Accelerated Evolution (AE) of convective simulations which
fast-forwards through the long thermal evolution of these systems, and we test the
validity of this method. The AE procedure involves measuring the dynamics of convection
early in a simulation and using its characteristics to adjust the mean thermodynamic
profile within the domain towards its evolved state. We study \RB convection as a test case for AE.  
Evolved flow properties of AE solutions are measured to be within a few percent
of solutions which are reached through Standard Evolution (SE) over a full thermal timescale.
At the highest values of the Rayleigh number at which we compare SE and AE,
we find that AE solutions require roughly an order of magnitude fewer cpu-hours
to compute than SE solutions.
\end{abstract}
\maketitle

@@

\section{Introduction}
\label{sec:intro}
Astrophysical convection occurs in the presence of disparate timescales. 
Studying realistic models of natural systems through direct numerical
simulations is infeasible due to the
large separation between various flow timescales and relaxation times.
Stiffness in astrophysical systems can manifest in multiple ways, some of which can
be handled by clever choices of numerical algorithms, and some which cannot.
For example,
flows in the convection zones of stars like the Sun are characteristically low Mach number
(Ma) in the deep interior. Initial value problems solved using
explicit timestepping methods are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit corresponding to the fastest motions (sound
waves), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff, and the difference between
the sound crossing time and the convective overturn time has made studies of low-Ma stellar
convection difficult. This stiffness can be avoided using approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out
\cite{brown&all2010, featherstone&hindman2016}.
Recently, advanced numerical techniques which use fully implicit 
\cite{viallet&all2011, viallet&all2013, viallet&all2016} or mixed
implicit-explicit \cite{lecoanet&all2014, anders&brown2017, bordwell&all2018} 
timestepping mechanisms have made it possible to study
convection in the fully compressible equations at low Mach numbers, 
and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

Unfortunately, astrophysical convective systems are stiff in more than one 
timescale. Specifically, the Peclet number (Pe), the ratio of the thermal
diffusion timescale to the convective velocity timescale, is large.
In a high Pe system, many convective timescales must pass before 
the domain thermally relaxes into a steady state in which
evolution of the thermal structure of the convective region has ceased.
As the timestep size of implicit methods is bound to the fastest nonlinear flows
(convection), fully implicit methods cannot be used to address this form of stiffness
\cite{viallet&all2011, viallet&all2013, viallet&all2016}.

Resolving dynamics in atmospheres which are sufficiently
thermally relaxed therefore remains a challenging problem.
Solar convection is a prime example of this phenomenon, as
dynamical timescales in the solar convective zone are relatively short 
(convection overturns every $\sim 5$ minutes at the solar surface)
compared to the Sun's thermal relaxation timescale, which is O($10^7$) years
\cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
meaningfully evolving the thermal structure of the system using
traditional timestepping techniques alone.
As modern simulations aim to model natural, high-Pe convection
in the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale ($t_{\kappa}$, defined in section \ref{sec:ae}) 
becomes intractably large compared to dynamical timescales 
such as the freefall time ($t_{\text{ff}}$, defined in section \ref{sec:experiment})
\cite{anders&brown2017}, 
\begin{equation}
\frac{t_{\kappa}}{t_{\text{ff}}} \propto (\text{Ra})^{1/2}.
\end{equation}
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Capturing appropriately resolved
turbulent motions requires finer grid meshes and smaller timesteps.
Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
and the number of overturn times required for systems to thermally relax
grows.

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally relaxed solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the flow
fields in a converged solution at low Ra as initial conditions for a simulation at high
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  In systems in which there are multiple
stable solutions, such as a roll state and a shearing state of convection,
the large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
It is possible that this puts the high Ra solution into a different stable state
than it would naturally reach from an initial, hydrostatically stable configuration.
Another commonly-used tactic in
moderate-Ra simulations is to first solve a simple model of the system in question,
and then use the solution of that model as initial conditions for full nonlinear
direct numerical simulations. For example, past studies have used initial conditions from
a linear eigenvalue solve \cite{hurlburt&all1984} in plane-parallel studies, or
axisymmetric solutions in studies of convection in 3D cylinders \cite{verzicco&camussi1997}. 
In other systems, particularly when convective zones are adjacent to stable regions,
authors often choose initial conditions which are not in a classic, hydrostatic,
conductive state. Rather, either through knowledge of low-Ra solutions \cite{couston&all2017}
or broader convective theories such as Mixing Length Theory
\cite{brandenburg&all2005}, initial conditions can be adjusted such that the stratification
within the convective domain is closer to a relaxed state than the largely
unstable hydrostatic state.

Despite the numerous methods that have been used,
the most straightforward way to achieve a thermally relaxed solution
is to evolve a convective simulation through a thermal timescale. Some modern
studies do just that \cite{featherstone&hindman2016}.
Such evolution is computationally 
\emph{expensive}, and state-of-the-art simulations at the highest values of Ra
can only reasonably be run
for hundreds of freefall timescales \cite{stevens&all2011}, much less the
thousands or millions of freefall times required for thermal relaxation.

In this work, we study a method of achieving accelerated evolution of
convective simulations. 
Our technique is similar in essence to the approach used in asymptotically 
reduced models of rapidly rotating convection, in which the
slowly evolving mean temperature profile is timestepped separately from the
fast convective dynamics \cite{julien&all1998, sprague&all2006}. 
We couple measurements of the dynamics of unstable, evolving
convective simulations with knowledge about energy balances in the desired solution
to instantaneously and self-consistently adjust the mean vertical thermodynamic 
profile toward its relaxed state. 
While a technique of this kind has been used by many studies previously 
(e.g., \cite{hurlburt&all1986}),
the details of implementation, the convergence properties, and whether or not the
thermally relaxed state achieved from 
accelerated evolution corresponds to the relaxed state in standard evolution 
are not documented.
In section \ref{sec:experiment}, we describe our convective simulations and
numerical methods. In section \ref{sec:ae}, we describe the procedure for 
achieving accelerated evolution. In
section \ref{sec:results}, we compare accelerated evolution solutions
to solutions obtained from standard evolution through a full thermal diffusion timescale. 
In section \ref{sec:speedups}, we compare the numerical cost of accelerated
solutions to standard solutions.
Finally,
in section \ref{sec:extensions}, we offer concluding remarks and
discuss extensions of the methods presented here.

@@

We study in depth a 2D simulation at $S = 10^5$ to demonstrate the power of AE.
We compare kinetic energy (KE, black line) and mean temperature (blue line)
traces from a SE run in Fig.~\ref{fig:time_trace}a to an AE run
in Fig.~\ref{fig:time_trace}c.
In Fig.~\ref{fig:time_trace}a, the time evolution of the SE simulation is shown.
The KE grows exponentially from white noise during the
first $\sim 25$ $t_{\text{ff}}$. The solution then saturates and begins to slowly
relax toward the saturated isothermal profile in the interior of the domain.
This slow relaxation is evident in the behavior of the blue line, which measures
$\angles{T} - T_{\text{top}}$, where \angles{T} is the volume-average of $T$, and
$T_{\text{top}} = -0.5$ is the temperature at the upper boundary.
The mean atmospheric temperature and
kinetic energies are fully converged when $t = 4000$ $t_{\text{ff}}$ $= 0.35$ $t_{\kappa}$.
We show roughly the first thousand freefall time
units of evolution, as well as the evolved thermodynamic state reached after a full
thermal time of evolution.  In Fig.~\ref{fig:time_trace}c, similar traces are
shown for the corresponding AE solution
at the same parameters. The same linear growth phase occurs, but shortly after
the peak of convective transient we accelerate the convergence of the atmosphere
through the process which we describe below. We adjust the 1D vertical profile of
the atmosphere three times, as denoted by the three labeled
arrows in the graph numbered 1-3.  The third profile adjustment associated with
arrow 3 is small enough (see appendix \ref{appendix:recipe}) that we assume
the atmosphere is sufficiently converged, and we begin to sample the evolved
convective dynamics.

@@

In order to adjust the temperature profile to achieve AE, we calculate the total flux,
F$_{\text{tot}} =$ F$_{\text{E}}$ + F$_{\kappa}$, and then derive the profiles
\vspace{-0.5cm}
\begin{equation}
f_{\text{E}}(z) = \frac{\text{F}_{\text{E}}}{\text{F}_{\text{tot}}},\qquad
f_{\kappa}(z) = \frac{\text{F}_{\kappa}}{\text{F}_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
which have the systematic asymmetries (Fig.~\ref{fig:time_trace}b) removed. These profiles describe which
parts of the atmosphere depend on convection to carry flux (where $f_{\text{E}}(z) = 1$
and $f_{\kappa}(z) = 0$).
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the extent of the early thermal boundary layers 
(where $f_{\kappa}(z) = 1$ and $f_{\text{E}}(z) = 0$) 
will not change significantly over the course of the atmosphere's evolution.
Under this assumption, in order to reach the converged state, 
the flux through the atmosphere must be decreased by some amount,
\vspace{-0.5cm}
\begin{equation}
\xi(z,t) \equiv \frac{\text{F}_{\text{B}}}{\text{F}_{\text{tot}}},
\label{eqn:xi}
\end{equation}
where $\text{F}_{\text{B}} = \mathcal{P}$ is the amount of flux that enters the
atmosphere at the bottom, fixed-flux boundary.
For example, in Fig.~\ref{fig:time_trace}b,
F$_{\text{tot}} \approx 19\mathcal{P}$ at the upper boundary,
but in the relaxed state 
it should just be $\mathcal{P}$, so $\xi \approx 1/19$ at that depth.

@@

\begin{figure}[b]
\includegraphics[width=\textwidth]{./figs/nu_v_time.png}
\caption{The time evolution of Nu (black) and $\angles{T} - T_{\text{top}}$ (blue)
are shown for simulations at $S = 10^5$ in SE (a) and AE (b).
A moving average of Nu, using a centered boxcar with a span of 50 freefall time
units, is overplotted as a grey line. 
We see that the mean value of Nu reaches its relaxed value quickly
compared to $\angles{T}$.
Regardless, fluctuations of Nu about the mean
value are much smaller in the thermally relaxed state.
The times at which AE solves occur are denoted by arrows in (b),
and the grey boxed region is examined in more detail in Fig. \ref{fig:oscillating_plumes}.
\label{fig:nu_v_time} }
\end{figure}

@@

The Nusselt number (Nu) quantifies the efficiency of convective heat transport
and is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}},
\end{equation}
where the volume average of a quantity, $\eta$, is shown as $\angles{\eta}$.
The time evolution of Nu in AE and SE is compared to the mean temperature evolution
in Fig.~\ref{fig:nu_v_time}. In Fig.~\ref{fig:nu_v_time}a, we show the evolution
of the SE run at $S = 10^5$. The black trace shows the instantaneous value of
Nu, and the overplotted grey line shows a moving time average of Nu. The
time average is taken using a centered boxcar window whose width is 50 $t_{\text{ff}}$.
The mean value of Nu evolves towards its final value (dashed horizontal
grey line) rapidly compared to $\angles{T}-T_{\text{top}}$ (blue), 
which approaches its evolved state (dashed horizontal blue line) slowly.
As such, it is possible to measure the statistically stationary mean value of
Nu after only a few hundred simulated freefall times, 
as has been done previously \cite{stevens&all2010}. However, fluctuations in
Nu around the mean value at early times have both a larger magnitude and higher
frequency than the fluctuations in the relaxed state.
In Fig.~\ref{fig:nu_v_time}b, we show the AE traces of Nu and
$\angles{T}-T_{\text{top}}$ at $S = 10^5$. The times at which AE solves occur are marked
by arrows. Nu reaches its evolved mean value slightly more rapidly than in the SE case,
and the frequency and magnitude of fluctuations in Nu away from the mean resemble the
final relaxed state of SE.

When $S < 10^{3.67}$ in 2D and for all runs in 3D, 
the evolved system is characterized by a time-stationary value of Nu, and is thus
in a state of constant convective heat transport.
At larger $S$ in 2D, the value of Nu varies significantly over time even in the
relaxed state (as seen in Fig.~\ref{fig:nu_v_time}). We examine the shaded
region of Fig.~\ref{fig:nu_v_time}b in more detail in the top left
panel of Fig.~\ref{fig:oscillating_plumes}, as well as a comparable time trace
at $S = 10^7$ (bottom, left panel). We find that these systems exhibit
large Nu during
states in which temperature fluctuations travel in their natural buoyant
direction (Fig.~\ref{fig:oscillating_plumes}, Ia \& IIa, where cold elements fall and hot elements rise).
However, when wrongly-signed temperature perturbations are entrained in an upflow or downflow
with oppositely signed fluid, Nu is suppressed (Fig.~\ref{fig:oscillating_plumes}, Ib and IIb, 
where warm fluid is pulled down by the downflow
lane, and cool fluid is drawn up by the upflow lane).
The plumes in these
systems naturally oscillate horizontally over time, switching between transport being dominated
by a counterclockwise cell, as pictured in Fig.~\ref{fig:oscillating_plumes}, and
a clockwise cell. Our choice of no-slip
boundary conditions prevents the fluid from entering a full domain shearing state 
\cite{goluskin&all2014}, and the
oscillatory motions of the plumes are a long-lived, stable phenomenon. 
However, thanks to our choice of periodic
boundary conditions and despite the no-slip conditions, the full system of the
upflow and downflow plumes is free to slowly migrate to the left or right over time,
and we observe this phenomenon in our solutions.
The 2D SE simulations exhibit the same horizontally
oscillatory behavior as the AE solutions for the same initial conditions. 
This time-dependent behavior of Nu is not seen strongly in our 3D solutions,
however most 3D simulations we conducted were at low $S$ compared to the runs in
which this behavior was observed in 2D.
